Constructing a universe for the setoid model

### 2020-10-12 paper.tex
Fix typos

The principle of propositional extensionality in the setoid model is an instance
of Voevodsky's univalence axiom \cite{hottbook}. The cubical set model is a
constructive model justifying this axiom \cite{cubical-sets}. A type theory
extracted from this model is Cubical Type Theory \cite{cubical}. The
relationship between the cubical set model and cubical type theory is similar to
that between the setoid model and SeTT.
%
Compared to cubical type theories, SeTT has the advantage that the equality type
satisfies more definitional equalities. For instance, whereas in cubical type
theory equality of functions is isomorphic to pair-wise equality, in SeTT the
isomorphism is replaced by a definitional equality. SeTT is also a syntactically
straightforward extension of Martin-L\"of Type Theory, that does not require
exotic objects like the internal pretype. In turn, the obvious advantage of
cubical type theory is that it is not limited to setoids.

@@

Types in a context $\GG$ are given by displayed setoids over $\GG$
with a fibration condition given by $\textsf{coe}$, $\textsf{coh}$. In
the following, we somtimes omit implicit quantifications such as the $\forall\{\gamma_0\ \gamma_1\}$ in the type of $\textsf{sym}\,\GG$.
\[
\infer{
  \begin{aligned}
    & | A | : | \GG | \to \mType \\
    & A^\sim : \{\gamma_0\ \gamma_1 : | \GG |\}
    \to \GG^\sim\ \gamma_0\ \gamma_1 \to | A | \gamma_0 \to | A | \gamma_1 \to \mProp \\
    & \textsf{refl*} : \{\gamma : | \GG | \} (a : | A | \gamma) \to A^\sim\ (\textsf{refl}\ \GG\ \gamma)\ a\ a \\
    & \textsf{sym*} : \forall\{\gamma_0\,\gamma_1\,a_0\,a_1\}\{p : \GG^\sim\ \gamma_0\ \gamma_1\}
    \to A^\sim\ p\ a_0\ a_1 \to A^\sim\ (\textsf{sym}\ \GG\ p)\ a_1\ a_0 \\
    & \textsf{trans*} : A^\sim\ p_0\ a_0\ a_1\to A^\sim\ p_1\ a_1\ a_2\to A^\sim\ (\textsf{trans}\,\GG\,p_0\,p_1)\ a_0\ a_2\\
    & \textsf{coe} : \GG^\sim\ \gamma_0\ \gamma_1 \to | A | \gamma_0 \to | A | \gamma_1 \\
    & \textsf{coh} : (p : \GG^\sim\ \gamma_0\ \gamma_1) (a : | A | \gamma_0)
    \to A^\sim\,p\,a\,(\textsf{coe}\,A\,p\,a) \\
  \end{aligned}
}{
  A : \Ty\ \GG
}
\]

@@

We can show that the setoid model validates the usual basic type formers ($\Pi,
\Sigma$, etc.), function extensionality and a universe of strict
propositions ith propositional extensionality \cite{setoid99}. Note that
we don't need identity types or inductive types (W-types) for this.

@@

We can show that $\Uty$ is closed under $\Pi$ types and booleans such
that $\EL\,(\mathsf{pi}\,A\,B) \equiv \Pi\,(\EL\,A)\,(\EL\,B)$ and
$\EL\,\mathsf{bool} = \mathsf{Bool}$. The universe can be closed under
more constructions if more codes are added to $\setoidU$.

>>>>

The principle of propositional extensionality in the setoid model is an instance
of Voevodsky's univalence axiom \cite{hottbook}. The cubical set model is a
constructive model justifying this axiom \cite{cubical-sets}. A type theory
extracted from this model is Cubical Type Theory \cite{cubical}. The
relationship between the cubical set model and cubical type theory is similar to
that between the setoid model and SeTT.
%
Compared to cubical type theories, SeTT has the advantage that the equality type
satisfies more definitional equalities. For instance, whereas in cubical type
theory equality of functions is isomorphic to pointwise equality, in SeTT the
isomorphism is replaced by a definitional equality. SeTT is also a syntactically
straightforward extension of Martin-L\"of Type Theory, that does not require
exotic objects like the interval pretype. In turn, the obvious advantage of
cubical type theory is that it is not limited to setoids.

@@

Types in a context $\GG$ are given by displayed setoids over $\GG$
with a fibration condition given by $\textsf{coe}$, $\textsf{coh}$. In
the following, we sometimes omit implicit quantifications such as the $\forall\{\gamma_0\ \gamma_1\}$ in the type of $\textsf{sym}\,\GG$.
\[
\infer{
  \begin{aligned}
    & | A | : | \GG | \to \mType \\
    & A^\sim : \{\gamma_0\ \gamma_1 : | \GG |\}
    \to \GG^\sim\ \gamma_0\ \gamma_1 \to | A | \gamma_0 \to | A | \gamma_1 \to \mProp \\
    & \textsf{refl*} : \{\gamma : | \GG | \} (a : | A | \gamma) \to A^\sim\ (\textsf{refl}\ \GG\ \gamma)\ a\ a \\
    & \textsf{sym*} : \forall\{\gamma_0\,\gamma_1\,a_0\,a_1\}\{p : \GG^\sim\ \gamma_0\ \gamma_1\}
    \to A^\sim\ p\ a_0\ a_1 \to A^\sim\ (\textsf{sym}\ \GG\ p)\ a_1\ a_0 \\
    & \textsf{trans*} : A^\sim\ p_0\ a_0\ a_1\to A^\sim\ p_1\ a_1\ a_2\to A^\sim\ (\textsf{trans}\,\GG\,p_0\,p_1)\ a_0\ a_2\\
    & \textsf{coe} : \GG^\sim\ \gamma_0\ \gamma_1 \to | A | \gamma_0 \to | A | \gamma_1 \\
    & \textsf{coh} : (p : \GG^\sim\ \gamma_0\ \gamma_1) (a : | A | \gamma_0)
    \to A^\sim\,p\,a\,(\textsf{coe}\,A\,p\,a) \\
  \end{aligned}
}{
  A : \Ty\ \GG
}
\]

@@

We can show that the setoid model validates the usual basic type formers ($\Pi,
\Sigma$, etc.), function extensionality and a universe of strict
propositions with propositional extensionality \cite{setoid99}. Note that
we don't need identity types or inductive types (W-types) for this.

@@

We can show that $\Uty$ is closed under $\Pi$ types and booleans, and satisfies
$\EL\,(\mathsf{pi}\,A\,B) \equiv \Pi\,(\EL\,A)\,(\EL\,B)$ and
$\EL\,\mathsf{bool} = \mathsf{Bool}$. The universe can be closed under more
constructions if more codes are added to $\setoidU$.
